MR-PFU: A suite of R packages for energy conversion chain analysis

### 2023-07-23 paper.Rmd
Add a table.

Energy flows through society enable economic activity and facilitate human flourishing.
To understand economic growth and human well-being, 
the field of energy analysis evaluates energy flows
from the primary stage 
(coal, oil, natural gas) 
to the final stage
(energy purchased by consumers, such as refined petroleum and electricity),
to the useful stage
(heat, mechanical drive, and light),
and sometimes to energy services
(transport, thermal comfort, and illumination).
Societal exergy analysis (SEA), an extension of energy analysis,
quantifies energy flows as exergy, 
the mechanical work potential of energy.

@@

Historically, SEA has analyzed energy flows through individual countries
using spreadsheets, 
with data in varying and inconsistent formats 
[@Ayres:2003ec; @Serrenho:2014aa; @Brockway:2014aa; @Brockway:2015aa].
The authors of this paper and others in the field
wanted to expand SEA to cover all countries, but
the spreadsheet approach to data analysis is not scalable.

To achieve scalable societal energy and exergy analysis, 
several sub-problems needed to be solved.
The first problem is the format for data storage. 
We use the Physical Supply-Use Table (PSUT) data structure,
a matrix approach to describing energy flows from 
extraction to
processing stages and, ultimately, 
to final demand [@Rocco:2016; @Guevara:2017; @Heun:2018].
A second problem is that `R` data fames.

It would be convenient to perform matrix mathematics in `R` data frames
as easily as scalar mathematics with the `dplyr` package from the `tidyverse`.
To do that, a way to embed matrices in a cell of a data frame is required.

>>>>

Energy flows through societal energy conversion chains (ECCs)
enable economic activity and facilitate human flourishing.
To understand economic growth and human well-being, 
the field of energy analysis evaluates ECCs
from the primary stage 
(coal, oil, natural gas) 
to the final stage
(energy purchased by consumers, such as refined petroleum and electricity),
to the useful stage
(heat, mechanical drive, and light),
and sometimes to energy services
(transport, thermal comfort, and illumination).
Societal exergy analysis (SEA), an extension of energy analysis,
quantifies ECCs as exergy, 
the mechanical work potential of energy.

@@

Historically, SEA has analyzed the ECCs of individual countries
using a large number of linked spreadsheets, 
with data in varying and inconsistent formats 
[@Ayres:2003ec; @Serrenho:2014aa; @Brockway:2014aa; @Brockway:2015aa].
The authors of this paper and others in the field
wanted to expand SEA to cover all countries, but
the spreadsheet approach to data analysis was
deemed not scalable.
We created a suite of `R` packages to provide scalable SEA 
for every country in the world.
This paper describes the design of those packages and
demonstrates briefly their use.

Several decisions influenced the design and APIs for the packages.
The decision involves data format. 
We selected the Physical Supply-Use Table (PSUT) data structure,
a matrix approach to describing energy flows from 
extraction to
processing stages and, ultimately, 
to final demand [@Rocco:2016; @Guevara:2017; @Heun:2018].
Given the decision to use PSUT matrices to store ECC data,
packages were created to enable manipulation of matrices
within the tidyverse, as shown in the following table.

| Package | Function |
|---------|----------|
| [RCLabels](https://github.com/MatthewHeun/RCLabels) | Manipulate row and column matrices |

It would be convenient to perform matrix mathematics in `R` data frames
as easily as scalar mathematics in data frames.

with the `dplyr` package from the `tidyverse`.
To do that, a way to embed matrices in a cell of a data frame is required.
